Triptych: logarithmic-sized linkable ring signatures with applications

### 2020-01-03 paper/iacr.tex
Paper updates

\documentclass{article}

\usepackage{enumerate}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[top=1.0in,bottom=1.0in,left=1.0in,right=1.0in]{geometry}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{defn}{Definition}[section]
\newtheorem{assu}{Assumption}[section]

\title{Compact linkable ring signatures and applications}
\author[1]{Brandon Goodell}
\author[1]{Sarang Noether}
\author[2]{\texttt{RandomRun}}
\affil[1]{Monero Research Lab, \texttt{\{surae,sarang\}@getmonero.org}}
\affil[2]{Independent researcher}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We describe an efficient linkable ring signature scheme, compact linkable spontaneous anonymous group (CLSAG) signatures, for use in confidential transactions. Compared to the existing signature scheme used in Monero, CLSAG signatures are both smaller and more efficient to generate and verify for ring sizes of interest. We generalize the construction and show how it can be used to produce signatures with coins of different type in the same transaction.
\end{abstract}

\section{Introduction}
Ring signatures are used in Monero to provide transaction signer ambiguity. Originally, a simple Schnorr-like linkable ring signature \cite{liu2004linkable} was used that included previous one-time output public keys as possible signers. Later, the addition of Pedersen commitments to hide amounts necessitated a generalization to the signature \cite{noether2016ring} in order to prove transaction balance without revealing the signer.

The current ring signature construction used in Monero, MLSAG signatures, scale linearly with the number of ring members included in the signature. In particular, they require the inclusion of two scalars per ring member.

We note that the MLSAG signature scheme, while general enough to permit linking across multiple sets of input keys, is in practice used as a hybrid approach to a linkable ring signature. Indeed, each signature input consists of two keys: the first is a standard one-time output public key, and the second is a key computed using the associated Pedersen commitments. Linkability is only used for the first key, to ensure that double-spending does not occur.

This construction is flexible. For example, the format can be used for usual ring confidential transactions or could be modified to describe \textit{colored ring confidential transactions} with multiple so-called coin colors, allowing users to transact with multiple assets simultaneously.

\subsection{Our contribution}

In this work, we introduce a $d$-dimensional linkable ring multisignature ($d$-LRMS) scheme suitable for use in a colored ring confidential transaction scheme with $d$ colors. Our scheme is \textit{compact} in the sense that signature size scales with the sum of ring size and the dimension $d$; that is, increasing $d$ yields the same additional number of signature elements, regardless of the ring size. We prove our scheme is unforgeable up to the hardness of the $k$-one-more discrete logarithm problem.

We call our signature scheme $d$-dimensional compact linkable spontaneous anonymous group ($d$-CLSAG) signatures. We demonstrate how to use this scheme for ring confidential transactions in Monero by replacing multi-layered linkable spontaneous anonymous group (MLSAG) signatures with $d$-CLSAG signatures.  The resulting scheme is more efficient in both signature size and verification time than current Monero transaction structures, and can be seamlessly integrated into the Monero transaction protocol.

\section{Notation}

In this document, we denote algorithms with typefont majuscule English letters like $\texttt{A}$, $\texttt{B}$, or $\texttt{O}$, or typefont names like $\texttt{Setup}$, $\texttt{KeyGen}$, and so on.

Group parameters are denoted as a tuple $(p, \mathbb{G}, d, G)$ where $\mathbb{G}$ is an elliptic curve group with prime order $p$, $d$ is a dimension, and $G$ is a generator of $\mathbb{G}$. We denote integers, bits, indices, and scalars in $\mathbb{Z}/p\mathbb{Z}$ with minuscule English letters $x, y, z, b, c, i, j, k$, etc.\ and we denote group elements with majuscule English letters, $G, X, W$, and so on. We use miniscule Greek letters like $\sigma$ to describe signatures and majuscule calligraphic Latin letters like $\mathfrak{T}$ when describing linkability tags.

We denote column vectors in boldface, e.g.\ $(x_1, \ldots, x_d)^\top = \textbf{x}$, and matrices in underlined boldface, e.g.\ $((x_{1,1}, x_{1,2}, \ldots, x_{1,n}), \ldots, (x_{d,1}, \ldots, x_{d,n})) = (\textbf{x}_1, \textbf{x}_2, \ldots, \textbf{x}_n) = \underline{\textbf{x}}$ is a $d\times n$ matrix. We denote the Hadamard product of two vectors with $\circ$, so for any $\textbf{x} = (x_1, x_2, \ldots, x_d) = (x_i)_{i=1}^{d}$, and for any $\textbf{y} = (y_i)_{i=1}^{d}$, we denote the sequence $(x_i \cdot y_i)_{i=1}^{d}$ with $\textbf{x} \circ \textbf{y}$. We denote bitwise concatenation with the symbol $\mid \mid$.

We distinguish oracles with calligraphic font, e.g.\ $\mathcal{CO}$ denotes a corruption oracle, $\mathcal{SO}$ denotes a signing oracle. If the codomain of a random oracle is the field of scalars $\mathbb{Z}/p\mathbb{Z}$, we denote this $\mathcal{H}^s$ (hash-to-scalar). If the codomain is $\mathbb{G}$, we denote this $\mathcal{H}^p$ (hash-to-point).

\section{Linkable ring multisignatures}

In this section, we recall linkable ring signature (LRS) schemes, we modify the definition to account for $d$-dimensional keys for linkable ring multisignature (LRMS) schemes, and we provide two examples. We emphasize that our description does not present multisignatures as computed between mutually untrusted parties. We use the term to mean a $d$-of-$d$ threshold multisignature with all keys controlled by the same signer; in particular, we have none of the communication steps necessary for computing a multisignature.

\subsection{LRS and LRMS schemes}

\begin{defn}[LRS]\label{def:lrs}
A \textit{linkable ring signature scheme} is a tuple $(\texttt{Setup}, \texttt{KeyGen}, \texttt{Sign}, \texttt{Verify}, \texttt{Link})$ satisfying the following. 
\begin{itemize}
\item $\texttt{Setup}(1^\lambda) \to \textit{par}$. $\texttt{Setup}$ takes as input a security parameter $1^\lambda$, produces some public parameters $\textit{par}$.

\item $\texttt{KeyGen}(1^\lambda, \textit{par}) \to (sk, pk)$. $\texttt{KeyGen}$ takes as input a security parameter $1^\lambda$ and public parameters $\textit{par}$. $\texttt{KeyGen}$ produces as output a private-public keypair $(sk, pk)$.

\item $\texttt{Sign}\left(1^\lambda, \textit{par}, (m, \textbf{pk}, sk)\right) \to \left\{\bot_{\texttt{Sign}}, (\sigma, \mathfrak{T})\right\}$. $\texttt{Sign}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, an arbitrary message $m \in \left\{0,1\right\}^*$, an ad hoc \textit{ring} of public keys $\textbf{pk} = \left\{pk_1, \ldots, pk_n\right\}$, and a secret key $sk$. $\texttt{Sign}$ produces as output either a distinguished failure symbol $\textit{out} = \bot_{\texttt{Sign}}$ or a signature-tag pair $\textit{out} = (\sigma, \mathfrak{T})$.

\item $\texttt{Verify}\left(1^\lambda, \textit{par}, (m, \textbf{pk}, (\sigma, \mathfrak{T}))\right) \to \left\{0,1\right\}$. $\texttt{Verify}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, a message $m$, a ring of public keys $\textbf{pk}$, and a signature-tag pair $(\sigma, \mathfrak{T})$. $\texttt{Verify}$ produces as output a bit $b \in \left\{0,1\right\}$.

\item $\texttt{Link}\left(1^\lambda, \textit{par}, (\sigma, \mathfrak{T}), (\sigma^\prime, \mathfrak{T}^\prime)\right)$. $\texttt{Link}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, and a pair of signature-tag pairs $(\sigma, \mathfrak{T}), (\sigma^\prime, \mathfrak{T}^\prime)$. $\texttt{Link}$ produces as output a bit $b \in \left\{0,1\right\}$. 
\end{itemize}
\end{defn}

We extend this to a scheme with $d$-dimensional keys.

\begin{defn}[$d$-LRMS]
A \textit{$d$-dimensional linkable ring multisignature scheme} is a tuple of algorithms $(\texttt{Setup}, \texttt{KeyGen}, \texttt{Sign}, \texttt{Verify}, \texttt{Link})$ satisfying the following. 
\begin{itemize}
\item $\texttt{Setup}(1^\lambda) \to \textit{par}$. $\texttt{Setup}$ works as described in \ref{def:lrs}.

\item $\texttt{KeyGen}(1^\lambda, \textit{par}) \to (\textbf{sk}, \textbf{pk})$. $\texttt{KeyGen}$ takes as input a security parameter $1^\lambda$ and public parameters $\textit{par}$. $\texttt{KeyGen}$ produces as output a private-public keypair $(\textbf{sk}, \textbf{pk})$ where each key is a $d$-dimensional vector. We refer to the first coordinate of each of these keys as the \textit{linking key} and the $d-1$ remaining coordinates (if there are any) are the \textit{auxiliary keys}.

\item $\texttt{Sign}\left(1^\lambda, \textit{par}, (m, \underline{\textbf{pk}}, \textbf{sk})\right) \to \left\{\bot_{\texttt{Sign}}, (\sigma, \mathfrak{T})\right\}$. $\texttt{Sign}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, a message $m$, a vector of public keys $\underline{\textbf{pk}} = \left\{\textbf{pk}_1, \ldots, \textbf{pk}_n\right\}$ where each $\textbf{pk}_i$ is a $d$-dimensional public key, and a $d$-dimensional secret key $\textbf{sk}$. $\texttt{Sign}$ produces as output either a distinguished failure symbol $\bot_{\texttt{Sign}}$ or a signature-tag pair $(\sigma, \mathfrak{T})$.

\item $\texttt{Verify}\left(1^\lambda, \textit{par}, (m, \underline{\textbf{pk}}, (\sigma, \mathfrak{T}))\right) \rightarrow \left\{0,1\right\}$. $\texttt{Verify}$ takes as input a security parameter $1^\lambda$, public parameters $\textit{par}$, a message $m$, a matrix of public keys $\underline{\textbf{pk}}$, and a signature-tag pair $(\sigma, \mathfrak{T})$. $\texttt{Verify}$ produces as output a bit $b \in \left\{0,1\right\}$.

\item  $\texttt{Link}\left(1^\lambda, \textit{par}, (\sigma, \mathfrak{T}), (\sigma^\prime, \mathfrak{T}^\prime)\right)$. $\texttt{Link}$ works as described in \ref{def:lrs}.
\end{itemize}
\end{defn}

We refer to the linkability tag as the \textit{key image} for consistency with terminology in Monero. We also use the following terminology if the scheme satisfies the respective properties.
\begin{description}
\item[Correctly verified:] For any message $m$, $d$-dimensional secret key $\textbf{sk}$ with corresponding public key $\textbf{pk}$, and appropriately-sized matrix $\underline{\textbf{pk}} = \left\{\textbf{pk}_1, \ldots, \textbf{pk}_n\right\}$, if there exists an index $1 \leq \ell \leq n$ such that $\textbf{pk} = \textbf{pk}_\ell$, then $\texttt{Verify}(\texttt{Sign}(m, \underline{\textbf{pk}}, \textbf{sk})) = 1$.

\item[Correctly linkable:] For any messages $m$, $m^\prime$, any $d$-dimensional secret keys  $\textbf{sk}$ and $\textbf{sk}^\prime$ with equal linking keys (that is, with $\textbf{sk}_0 = \textbf{sk}^\prime_0$) and with corresponding public keys $\textbf{pk}$ and $\textbf{pk}^\prime$ respectively, any appropriately-sized pair of matrices $\underline{\textbf{pk}}$ and $\underline{\textbf{pk}}^\prime$, if $(\sigma, \mathfrak{T}) \leftarrow \texttt{Sign}(m, \underline{\textbf{pk}}, \textbf{sk})$ and $(\sigma^\prime, \mathfrak{T}^\prime) \leftarrow \texttt{Sign}(m^\prime, \underline{\textbf{pk}}^\prime, \textbf{sk}^\prime)$ and there exists a pair of indices $\ell$ and $\ell^\prime$ such that $\underline{\textbf{pk}}_\ell = \textbf{pk}$ and $\underline{\textbf{pk}}_{\ell^\prime} = \textbf{pk}^\prime$, then
$\texttt{Link}((\sigma, \mathfrak{T}), (\sigma^\prime, \mathfrak{T}^\prime))=1$.
\end{description}

Note that linkability is only considered with respect to a single component of a matrix entry. When applied to Monero, this is important since only one such entry corresponds to an output public key for double-spend detection purposes.

%In the sequel, we omit $1^\lambda$ and $\textit{par}$ from the input of algorithms for clarity unless there is risk of confusion.

\subsection{Examples}

\begin{example}\label{ex:lsag}
The LSAG signature scheme from \cite{liu2004linkable} is a $1$-LRMS. In this example, $\texttt{Setup}$ always deterministically sets $d:=1$ (so we only use the linking key and there are no auxiliary keys). $\texttt{Setup}$ selects some $G \in \mathbb{G}$ to be a group generator for the group parameters $(p, \mathbb{G}, 1, G)$, two cryptographic hash functions $\mathcal{H}^s:\left\{0,1\right\}^* \to \mathbb{Z}/p\mathbb{Z}$ and $\mathcal{H}^p:\left\{0,1\right\}^* \to \mathbb{G}$. $\texttt{Setup}$ outputs $\textit{par} = (p, \mathbb{G}, 1, G, \mathcal{H}^s, \mathcal{H}^p)$.

$\texttt{KeyGen}$ produces as output a private key $x \in \mathbb{Z}/p\mathbb{Z}$ and a public key $X = xG$. $\texttt{Sign}$ takes as input a private key $y \in \mathbb{Z}/p\mathbb{Z}$, a message $m$, and a ring $\textbf{X} = \left\{X_1, \ldots, X_n\right\}$, and produces as output either a distinguished failure symbol $\bot_{\texttt{Sign}}$ or a signature-tag pair $(\sigma, \mathfrak{T})$.

The signature scheme originally described in \cite{liu2004linkable} signs a message $m$ with a ring of keys $\textbf{X} = \left\{X_1, \ldots, X_n\right\}$ and a secret key $x$ corresponding to some $X_\ell$, using the linkability tag (key image) $\mathfrak{T} := x \mathcal{H}^p(\textbf{X})$. Unfortunately, this key image is unsuitable for use in Monero, as changing ring members will change the key image. This would allow the same key to be used twice in two different ring signatures. For use in Monero, we modify this key image to be independent of the non-signing ring members, $\mathfrak{T} := x \mathcal{H}^p(X_\ell)$. This allows these key images to be used for double-spend protection.

An LSAG signature on a message $m$ with ring $\textbf{X}$ using these key images is computed in the following way. First, the signer samples $\alpha, s_{\ell+1}, s_{\ell+2}, \ldots, s_{\ell-1} \in \mathbb{Z}/p\mathbb{Z}$ at random. Next, the signer computes basepoints $H_i = \mathcal{H}^p(X_i)$. Next, the signer computes $c_{\ell+1}  = \mathcal{H}^s(\textbf{X} \mid \mid m \mid \mid \alpha G \mid \mid \alpha H_\ell)$ and the signer computes each $c_{i+1} = \mathcal{H}^s(\textbf{X} \mid \mid m \mid \mid s_i G + c_i X_i \mid \mid s_i H_i + c_i \mathfrak{T})$ for $i = \ell+1, \ldots, \ell-1$, naturally identifying index $1$ with index $n+1$. The signer finishes by computing $s_\ell = \alpha - c_\ell x_\ell$ and publishing the signature-tag pair $(\sigma, \mathfrak{T})$ where $\sigma = (c_1, s_1, \ldots, s_n)$.

A purported LSAG signature-tag pair on a message $m$ with ring $\textbf{X}$ is verified in the following way. The verifier sets $c_1^\prime = c_1$ and computes $c_{i+1}^\prime = \mathcal{H}^s(\textbf{X} \mid \mid m \mid \mid s_i G + c_i^\prime X_i \mid \mid s_i H_i + c_i^\prime \mathfrak{T})$ for $i=1, 2, \ldots, n$. The verifier outputs $1$ when $c_{n+1}^\prime = c_1$ and $0$ otherwise. Note that all verifiers must list the keys in $\textbf{X}$ in an agreed-upon order for the above verification to work; either they should agree upon lexicographic or some other ordering.  LSAG signature-tag pairs are linked merely by comparing key images: two valid signatures with the same key image were signed with the same secret key (and, in the case of Monero, would signal an attempt to double-spend funds).
\end{example}

\begin{remark}
The key image modification in Example \ref{ex:lsag} is due to the basepoint of the key image $\mathfrak{T}$. As noted in \cite{liu2004linkable}, easy variations on key image formats are available. How or whether the security properties of LSAG signatures are retained in practical use given more flexible key image formats, while interesting, is beyond the scope of this work.
\end{remark}

>>>>

\documentclass[draft]{article} % TODO: remove draft
\usepackage[top=1.0in,bottom=1.0in,left=1.0in,right=1.0in]{geometry}
\usepackage{amsmath,amssymb,amsthm,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{graphicx}

% TODO remove
\usepackage{lineno}
\linenumbers

\title{Triptych: logarithmic-sized linkable ring signatures with applications}
\author{Sarang Noether \\ Monero Research Lab \\ \texttt{sarang.noether@protonmail.com}}
\date{\today}

\newcommand{\G}{\mathbb{G}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\com}{\operatorname{Com}}
\newcommand{\hs}{\mathcal{H}}
\newcommand{\hp}{\mathbb{H}}

\newcommand{\A}{\mathcal{A}}

\newcommand{\sumi}{\sum_{i=0}^{n-1}}
\newcommand{\sumj}{\sum_{j=0}^{m-1}}
\newcommand{\sumk}{\sum_{k=0}^{N-1}}

\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

\begin{document}

\maketitle

\begin{abstract}
Ring signatures are a common construction used to provide signer ambiguity among a non-interactive set of public keys specified at the time of signing.
Unlike early approaches where signature size is linear in the size of the signer anonymity set, current optimal solutions either require centralized trusted setups or produce signatures logarithmic in size.
However, few also provide linkability, a property used to determine whether the signer of a message has signed any previous message, possibly with restrictions on the anonymity set choice.
Here we introduce Triptych, a family of linkable ring signatures without trusted setup that is based on generalizations of zero-knowledge proofs of knowledge of commitment openings to zero.
We demonstrate applications of Triptych in signer-ambiguous transaction protocols by extending the construction to openings of parallel commitments in independent anonymity sets.
Signatures are logarithmic in the anonymity set size and, while verification complexity is linear, collections of proofs can be efficiently verified in batches.
We show that for anonymity set sizes practical for use in distributed protocols, Triptych offers competitive performance with a straightforward construction.
\end{abstract}

\section{Introduction}
First introduced in \cite{rivest} with respect to RSA groups, ring signatures permit the signing of messages using public key sets not fixed in advance, without the need of a trusted group manager.
Earlier constructions lacked such flexibility, requiring either centralized key setup or the establishment of fixed signing sets.
Later work \cite{bender} established more robust security models for unforgeability and anonymity, capturing realistic threat models where an adversary is permitted to corrupt keys, convince honest signers to include malicious anonymity set members, or obtain signatures in advance.

Since a ring signature has an anonymity set of public keys, one of which is the true signer, the detection of signing by the same key requires an additional property, linkability.
A linkable ring signature \cite{lsag} enables verifiers to determine whether the (unknown) signer of a message has signed other messages.
Such a construction was proposed for election applications, where it is necessary to ensure that votes are anonymous, but voters are permitted to vote only once on a particular issue.
The construction in \cite{lsag}, with a hash-trapdoor structure similar to that of \cite{schnorr}, is of particular interest due to potential flexibility in linking; while its linking is limited to signer-selected but unchanging signer anonymity sets, it permits linking on a per-issue basis.
Additional recent work in \cite{backes} introduces the property of linkable anonymity, using sets of signatures and establishing restrictions on key corruption.
Other related interesting properties like traceability \cite{fs,f} imply stronger capabilities, where an attempt to sign two messages with the same key permits verifiers to identify the signer.

Linkable ring signatures have seen particular applications in signer-ambiguous transaction protocols.
In such an application, transactions are authorized by a ring signature whose signing anonymity set consists of previously-generated transaction outputs.
A signature demonstrates that the signer controls the private key of one such output without revealing which is the signer, and linkability is used to assure verifiers that output has not been used in another signature (signifying a double-spend attempt).

A practical consideration for the use of linkable ring signatures in transaction protocols is how signature size and verification time scale with the size of the anonymity set.
In common applied constructions like \cite{mlsag,clsag}, signature size and verification time scale linearly with the size of the signing ambiguity set; since such signatures are typically included in a public distributed data structure like a blockchain, there is a balance between the size of the anonymity set and the requirements for storage and verification.
Recent protocol work in this area mitigates the size restriction.
For example, in \cite{rct3}, the authors introduce a confidential transaction protocol based on a proving system whose size scales logarithmically with the anonymity set size, and which includes a method for demonstrating amount balance; amount commitment range proving is offloaded to other constructions like \cite{bulletproofs}.
In \cite{omniring}, the authors use a more general proving technique to accomplish a similar goal; however, the protocol offers further size benefits by integrating commitment range proofs into the proof structure directly, taking advantage of the logarithmic proof size.

Other signer-ambiguous transaction protocols not based on linkable ring signatures offer more competitive performance.
For example, protocols like \cite{zcash} produce maximal theoretical signer ambiguity through the use of zero-knowledge Merkle proofs (among others) that offer extremely small proofs with low verification time, but at the cost of a trusted structured setup process that arises from the underlying proving system \cite{groth16}.
Like this work, \cite{lelantus} is a transaction protocol also based on \cite{groth}; however, it is intended to operate on commitments similar to those used in \cite{zerocoin} but with the inclusion of amounts, and has limitations on addressing and sender tracing.

\subsection{Our contribution}
We produce two instantiations of a linkable ring signature, which we call Triptych.
The constructions are a linkable generalization of Groth's one-of-many commitment-to-zero proving system \cite{groth}, with optimizations from Bootle \cite{bootle} applied for improvements to proof size and verification complexity.
In one version of Triptych, the prover shows that it knows the opening of a commitment to zero within a commitment list, and also that it has constructed a linking tag using the same opening, enabling linkability.
We then modify Groth's ring signature definitions to include linkability and a related property, non-frameability.

In the second version of Triptych, we further generalize to include a parallel list of commitments.
Here, the prover proceeds as before to show its knowledge of an opening to a commitment in one list, as well as the construction of the linking tag.
However, the proof also shows that the prover knows an opening of a commitment at the same position in the second commitment list.
This construction has immediate application; in some signer-ambiguous confidential transaction protocols, transaction inputs are commitments to zero, for which the signer shows it knows the opening.
Each commitment comes equipped with another commitment to the amount of the input; by offsetting these commitments homomorphically and carefully choosing commitment randomness, the prover can show that a particular transaction balances.

We show that Triptych produces signatures with competitive performance to other modern linkable ring signatures for limited anonymity set sizes.
We note that similar constructions also require linear verification time, meaning that the size of anonymity set used in practice is likely to be limited for performance reasons.

\section{Preliminaries}
\subsection{Public parameters}
Let $\G$ be a cyclic group in which the discrete logarithm problem is hard, and let $\F$ be the scalar field of $\G$.
Let $\hs: \{0,1\}^* \to \F$ be a cryptographic hash function.
Let $G$ and $H$ be generators of $\G$ with unknown discrete logarithm relationship.
Let $N = n^m$ be a size parameter, where $n > 1$ and $m > 1$.
Let $\{G_{j,i}\}_{j,i=0}^{m-1,n-1}$ be a set of generators of $\G$ with unknown discrete logarithm relationship to each other, to $G$, and to $H$.
Let $U$ be a generator of $\G$.
Note that all generators may be produced using public randomness; for example, the use of a suitable hash function with domain separation may be appropriate.
All such public parameters are assumed to comprise a global reference string known to all players; in particular, we exclude them from algorithm definitions and Fiat-Shamir transcript hashes for readability.

\subsection{Pedersen commitment}
Let $\com$ be a homomorphic commitment scheme that is perfectly hiding and computationally binding.
In this work, we assume use of the Pedersen commitment scheme: for $x,r \in \F$, define $\com(x,r) \equiv xG + rH$ to be the commitment of the value $x$ with randomness $r$.
This can be trivially extended to support matrix values; for $\{x_{j,i}\},r \in \F$, define $\com(x,r) \equiv rH + \sum_{j,i} x_{j,i}G_{j,i}$.
Note in particular that Pedersen matrix commitments are similarly homomorphic.

\subsection{Other notation}
For integers or field elements $i,j$, the Kronecker delta function $\delta(i,j)$ evaluates to $1$ if $i=j$ and $0$ otherwise, where the output is taken to be in the appropriate set.

We sometimes use index subscript notation of the form $i_j$ to indicate the $j$ digit of $i$, where such a decomposition of $i$ is taken base $n$ with padded length $m$:
$$\sum_{j=0}^m i_j n^j = i$$
This notation will be specified explicitly where confusion may occur.

\section{Protocol: linkable one-of-many commitment}
We wish to build a linkable ring signature construction, where a signer who knows the opening of a commitment may sign messages using an anonymity set containing other commitments for which the signer does not know openings.
Included with the proof of knowledge, the signer also provides a linking tag that is the image of the signing commitment's opening under a verifiable pseudorandom function, using the method of \cite{dodis} that has previously appeared in \cite{omniring,rct3}.
Part of the soundness of the proving system relies on the proper construction of this linking tag.
Upon receipt, a verifier can check whether the linking tag has previously appeared in any other valid proof; if it has not, injectivity assures the verifier that no other signature has been produced by the (unknown) signer.

More specifically, we modify the construction of Bootle \cite{bootle}, which itself is a generalization of a construction by Groth \cite{groth}.
We produce a sigma protocol for the following relation:
$$\mathcal{R}_{\text{link}} = \left\{ \{M_i\}_{i=0}^{N-1} \subset \G, J \in \G ; (l \in \mathbb{Z},r \in \F) : M_l = rG \text{ and } U = rJ \right\}$$
Figures \ref{fig:link} and \ref{fig:link_2} describe the protocol.

\begin{figure}[htbp]
\centering
\fbox{\begin{minipage}{0.95\textwidth}
$\underline{\mathcal{P}_{\text{link}}(\{M_i\},J;(l,r)):}$
\begin{itemize}
\item Select random $r_A \in \F$ and $\left\{a_{j,i}\right\}_{i=1,j=0}^{n-1,m-1} \subset \F$.
Define $A \equiv \com(a,r_A)$.
\item Define $\left\{\sigma_{j,i}\right\}_{i,j=0}^{n-1,j-1} \subset \F$ such that $\sigma_{j,i} \equiv \delta\left(l_j,i\right)$ (using our decomposition notation), and choose random $r_B \in \F$.
Define $B \equiv \com(\sigma,r_B)$.
\item Select random $r_C \in \F$, and define $C \equiv \com(a(1-2\sigma), r_C)$.
\item Select random $r_D \in \F$, and define $D \equiv \com(-a^2, r_D)$.
\item Define coefficients $\left\{p_{k,j}\right\}_{k,j=0}^{N-1,m-1}$ such that $$p_k(x) \equiv \prod_{j=0}^{m-1} \left( \sigma_{j,k}x + a_{j,k} \right) = \delta\left(l,k\right)x^m + \sumj p_{k,j}x^j$$ for all $k \in [0,N)$ (using our decomposition notation).
\item Select random $\left\{\rho_j\right\}_{j=0}^{m-1} \subset \F$.
\item Define $\{X_j\}_{j=0}^{m-1} \subset \G$ such that: $$X_j \equiv \sumk p_{k,j}M_k + \rho_jG$$
\item Define $\{Y_j\}_{j=0}^{m-1} \subset \G$ such that: $$Y_j \equiv U \sumk p_{k,j}+ \rho_jJ$$
\end{itemize}

$\underline{\mathcal{P} \to \mathcal{V}:}$ \\
$A,B,C,D,\{X_j\},\{Y_j\}$ \\

$\underline{\mathcal{V} \to \mathcal{P}:}$ \\
$\xi \in \{0,1\}^*$ \\

$\underline{\mathcal{P}(\xi):}$
\begin{itemize}
\item Define $\left\{f_{j,i}\right\}_{i=1,j}^{n-1,m-1}$ such that $f_{j,i} \equiv \sigma_{j,i}\xi + a_{j,i}$.
\item Define $z_A \equiv r_A + \xi r_B$ and $z_C \equiv \xi r_C + r_D$.
\item Define $z \equiv r \xi^m - \sumj \rho_j \xi^j$.
\end{itemize}

$\underline{\mathcal{P} \to \mathcal{V}:}$ \\
$\{f_{j,i}\}_{j=0,i=1}^{m-1,n-1},z_A,z_C,z$
\end{minipage}}
\caption{Sigma protocol for $\mathcal{R}_{\text{link}}$}
\label{fig:link}
\end{figure}

\begin{figure}[htbp]
\centering
\fbox{\begin{minipage}{0.95\textwidth}
$\underline{\mathcal{V}_{\text{link}}(\{M_i\},J):}$
\begin{itemize}
\item For $0 \leq j < m$, let $f_{j,0} \equiv \xi - \sum_{i=1}^{n-1} f_{j,i}$.
\item Accept if and only if:
\begin{eqnarray}
A + \xi B &=& \com(f,z_A) \label{eqn:ab} \\
\xi C + D &=& \com(f(\xi-f),z_C) \label{eqn:cd} \\
\sumk M_k \left( \prod_{j=0}^{m-1} f_{j,k_j} \right) - \sumj \xi^jX_j - zG &=& 0 \label{eqn:x} \\
U \sumk \left( \prod_{j=0}^{m-1} f_{j,k_j} \right) - \sumj \xi^jY_j - zJ &=& 0 \label{eqn:y}
\end{eqnarray}
\end{itemize}
\end{minipage}}
\caption{Sigma protocol for $\mathcal{R}_{\text{link}}$ (continued)}
\label{fig:link_2}
\end{figure}

Observe that this protocol can be made non-interactive using the Fiat-Shamir heuristic, where the verifier challenge is produced using a collision-resistant hash function (modeling a random oracle) and the proof transcript \cite{fiatshamir}.

We will show that the sigma protocol is complete, sound, and zero-knowledge, the precise definitions of which are common and found in \cite{groth}.
Informally, we require the protocol be:
\begin{itemize}
\item \textit{Perfectly complete}: Given knowledge of a witness to a statement in the proof relation, an honest prover can always convince an honest verifier of the validity of the witness.
\item \textit{Special sound}: Given a statement in the proof relation, if a prover can answer multiple honest verifier challenges correctly, then it is possible to extract a witness for this statement.
\item \textit{Special honest-verifier zero knowledge}: Given any statement and verifier challenge, it is possible to simulate a transcript that is accepted by an honest verifier without knowledge of a corresponding witness.
\end{itemize}

\begin{theorem}
The protocol in Figures \ref{fig:link} and \ref{fig:link_2} is perfectly complete, special honest-verifier zero knowledge, and $(m+1)$-special sound.
\end{theorem}

\begin{proof}
The proof follows similarly to that of \cite{bootle}.

% Completeness
We first show perfect completeness.
Suppose the verifier receives a proof generated by an honest prover.
Equation \ref{eqn:ab} holds using the identity $$\sumi \sigma_{j,i} = 1$$ for all $0 \leq j < m$.
Equation \ref{eqn:cd} follows similarly, using the identity $$\left(\sigma_{j,i}\right)^2 = \sigma_{j,i}$$ for all $0 \leq j < m$.
To show Equation \ref{eqn:x} holds:
\begin{eqnarray*}
&& \sumk M_k \left( \prod_{j=0}^{m-1} f_{j,k_j} \right) - \sumj \xi^jX_j - zG \\
&=& \sumk M_k p_k(\xi) - \sumj \xi^j \left( \sumk p_{k,j}M_k + \rho_jG \right) - zG \\
&=& \sumk M_k \left( p_k(\xi) - \sumj \xi^j p_{k,j} \right) - \sumj \xi^j \rho_jG - zG \\
&=& \sumk M_k \xi^m \delta(l,k) - \sumj \xi^j \rho_jG - \left( r\xi^m - \sumj \rho_j\xi^j \right)G \\
&=& \xi^mrG - \sumj \xi^j \rho_jG - \xi^mrG + \sumj \xi^j \rho_jG \\
&=& 0
\end{eqnarray*}
Equation \ref{eqn:y} follows similarly:
\begin{eqnarray*}
&& U \sumk \left( \prod_{j=0}^{m-1} f_{j,k_j} \right) - \sumj \xi^jY_j - zJ \\
&=& U \sumk p_k(\xi) - \sumj \xi^j \left( U \sumk p_{k,j} + \rho_jJ \right) - zJ \\
&=& U \sumk \left( p_k(\xi) - \sumj \xi^j p_{k,j} \right) - \sumj \xi^j \rho_jJ - zJ \\
&=& U \sumk \xi^m \delta(l,k) - \sumj \xi^j \rho_jJ - \left( r\xi^m - \sumj \rho_j\xi^j \right)J \\
&=& \xi^mU - \sumj \xi^j \rho_jG - \xi^mrJ + \sumj \xi^j \rho_jG \\
&=& 0
\end{eqnarray*}
since $J = r^{-1}U$ in a valid proof.
Hence the protocol is perfectly complete.

% SHVZK
We next show that the protocol is special honest-verifier zero knowledge.
To do so, we construct a simulator that, given a random verifier challenge $\xi$, can construct a proof transcript with identical distribution to a valid proof.

First, observe that the simulator presented in the proof of Lemma 1 in \cite{bootle} translates identically to our setting.
If the simulator chooses $B \in \G$ uniformly at random, the cited lemma assures us a valid simulation of the proof elements $A,C,D,z_A,z_C,\{f^{(u)}_{j,i \neq 0}\}$; we may compute each $f^{(u)}_{j,0}$ from this.
Further, in a valid proof, $B$ is independent and uniformly distributed as well.

The proof elements $\{X_j\}_{j=1}^{m-1}$ and $\{Y_j\}_{j=1}^{m-1}$ are independent and uniformly distributed in a valid proof since the set $\{\rho_j\}$ is random and the discrete logarithm problem in $\G$ is hard, so the simulator may choose these uniformly at random.
The verification checks require that $X_0$ and $Y_0$ be uniquely determined by the other elements in the corresponding sets in both real proofs and by the simulator.

Finally, $z$ is uniformly distributed in valid proofs given random $\xi$, so the simulator may choose it uniformly at random.
Hence the construction is special honest verifier zero-knowledge.

% Soundness
It remains to show that the protocol is $(m+1)$-special sound, where $m > 1$.
To show this, we construct an extractor that, given $m+1$ valid responses to $m+1$ distinct verifier challenges for the same initial statement, produces a valid witness.
